Load, Clean, and Save an SPSS File
========================================================

Load the raw data file from the SPSS format, clean the data, and save it to CSV

### Inputs
* A single file in SPSS format

### Outputs
* One cleaned version of the data, in comma-separated-variable format
* One cleaned version of the data, in R format

### Configuration Settings (set in the code block below)
* Working directory path
* File name of the SPSS file
* File name for the cleaned and converted output files
* (Optional) A list of column numbers to remove from the data

The R format data file can be loaded by R for future analysis using the command:
```
interesting_data <- readRDS("filename.R")
```
where:
* *interesting_data* is your variable name for the dataframe
* *filename.R* is the filename you specified (below) for **clean_file_name**

## Set project parameters
```{r}
working_directory = "~"
raw_file_name = "your_research_file.sav"

# multiple versions of the clean file will be saved, and the appropriate extension will be applied
clean_file_name = "Your research file name"
```

## Load Raw Data
```{r}
library(foreign)
setwd(working_directory)
raw_data <- read.spss(raw_file_name, to.data.frame = TRUE)
cat("Sample size =",dim(raw_data)[1],"for",dim(raw_data)[2],"questions / variables")
colnames(raw_data)
```

## Clean Data
### Identify unnecessary columns (automatically)
Automatically select any columns for removal that have any of the following characteristics:
* no data
* the data in every row is the same

```{r}
cut_list <- 0
# Iterate across all columns in the data set
for(col_num in 1:dim(raw_data)[2]){
  this_column <- raw_data[,col_num]

  # Different tests for factors and other responses
  if(is.factor(this_column)){
    
    # Only one level means all the data is the same, and therefore not interesting
    if(length(levels(this_column))<2){
      cut_list <- c(cut_list,col_num)
    }else{
      #Need to determine if a factor with 2 or more levels is using more than one of them
      if(max(table(this_column)) == length(this_column)){
        cut_list <- c(cut_list,col_num)
      }
    }
  }else{
    # These columns are not factors
    if(length(table(this_column))<2){
      # All rows have the same value, so this column is not interesting
      cut_list <- c(cut_list,col_num)
    }
  }
}

cut_list <- cut_list[2:length(cut_list)]
```
### Highlight columns you may want to remove
```{r}
cut_list
```

### Clear out unnecessary columns (manually)
Remove any columns that are blank, that were meta-data / not part of the questionaire, or that all have the same answer
This list should be edited manually; possibly with guidance from the automatically generated list above.
```{r}
# put any column numbers you wish to remove between the paranthesis below
# E.g. remove <- c(1,2,3,4)
# You can specify a range of columns to remove by using a colon, like this c(2:4, 10:25, 30, 32)
remove <- c()

clean_data <- subset(raw_data,select=-remove)
cat("Sample size =",dim(clean_data)[1],"for",dim(clean_data)[2],"questions / variables")
colnames(clean_data)
```

## Save to CSV file and to R object file

```{r}
write.csv(clean_data, file = paste(clean_file_name,"csv", sep="."), row.names = FALSE)
saveRDS(clean_data, file = paste(clean_file_name,"R", sep="."))
```
